\documentclass[letterpaper]{article}

\usepackage{fullpage}
\usepackage{nopageno}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage[utf8]{luainputenc}
\usepackage{aeguill}
\usepackage{setspace}

\tikzstyle{edge} = [fill,opacity=.5,fill opacity=.5,line cap=round, line join=round, line width=50pt]
\usetikzlibrary{graphs,graphdrawing}
\usegdlibrary{trees}

\pgfdeclarelayer{background}
\pgfsetlayers{background,main}

\allowdisplaybreaks

\newcommand{\abs}[1]{\left\lvert #1 \right\rvert}

\begin{document}
\title{Notes}
\date{13 mars, 2015}
\maketitle
\section*{the happy ending problem}
\subsubsection*{task:}
draw 5 points in a plane with no 3 colinear points.

this is called general position.

\subsubsection*{question}
are there 4 points that form a convex quadrilateral?

\subsubsection*{answer}
yes

\subsection*{proof}
only way to have a non convex quadrilateral is to have a triangle with a dot in the middle. the options for where to put the fifth dot are: in the concave bit, which forms one, or in the pennisula in the quad, next to the concave bit, and so on

at nine points you can always find a convex pentagon

what is the minimum number of points in general required to guarantee a convex n-gon?

3 for 3, 4 requires 5, 5 requires 9, $C(6)=17$ the general answer is unknown, but $C(n)\ge 1+2^{n-2}$ where $C(n)$ is the number of points and $n$ is the order of the convex n-gon. this theorem is from forties or fifties.

\subsubsection*{thrm}
$C(n)\le \binom{2n-5}{n-2}<\infty$

above 6 the problem is open.


\section*{ramsey theory}
what is the minimum number of widgets with proposition required to guarantee another proposition?

\subsection*{modification}
require no points inside convex polygon

then $C'(3)=3, C'(4)=5, C'(5)=?$

\subsubsection*{thrm}
if only empty n-gons are allowed, then $C(7)\not<\infty$
\end{document}
