\documentclass[letterpaper]{article}

\usepackage{fullpage}
\usepackage{nopageno}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tikz}
\usepackage[utf8]{luainputenc}
\usepackage{aeguill}
\usepackage{setspace}

\tikzstyle{edge} = [fill,opacity=.5,fill opacity=.5,line cap=round, line join=round, line width=50pt]
\usetikzlibrary{graphs,graphdrawing}
\usegdlibrary{trees}

\pgfdeclarelayer{background}
\pgfsetlayers{background,main}

\allowdisplaybreaks

\newcommand{\abs}[1]{\left\lvert #1 \right\rvert}

\begin{document}
\title{Notes}
\date{13 avril, 2015}
\maketitle
8.4 list colorings,
9.1 4 color theorem

recall a list coloring is a proper coloring of $G$ ($C:V(G)\to\mathbb{Z}$) where each vertex has a restricted list of possible colors

\section*{4 color thm}
every planar graph is 4 colorable, ie $G$ planar implies chromatic G less than 4 $\chi(G)\le 4$

\section*{9.2 colorings of planar graphs}

n friday we mentioned Haewood's counterexample to part of kempe's (erroneous) proof of the 4 color theorem.

\subsection*{haewood 5 color theorem}
$G$ planar implies $\chi(G)\le 5$.

\subsubsection*{proof} in book, like half a page.

\subsection*{2 questions}
\subsubsection*{what planar graphs have chromatic number less than 4?}
\begin{enumerate}
\item
disconnected points (trivial graphs)
\item
bipartite (including trees)
\item
when does this happen?

partial answer:
a maximal planar graph $G$ of order 3 or more has chromatic number 3 if and only if $G$ is Eulerian.

you can remove edges without making coloring go up?
\end{enumerate}
\subsubsection*{is every planar graph 4-choosable}


\subsection*{thomassen thm}
every planar graph is 5-choosable

are there planar graphs that are not 4-choosable? answer is yes.

\subsubsection*{thm}
Mirzakhani graph is planar but not 4-choosable and has order 63 which betters the 238 order previous counter example.

homework:
3,4 due with 8.5
\end{document}
