\documentclass[letterpaper]{article}

\usepackage{fullpage}
\usepackage{nopageno}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{aeguill}
\usepackage{setspace}
\usepackage{apacite}
\allowdisplaybreaks

\newcommand{\abs}[1]{\left\lvert #1 \right\rvert}

\begin{document}
\title{Controversy Outline\\Persuasive Presentation}
\date{April 17, 2015}
\author{Jon Allen}
\maketitle
\renewcommand{\labelenumi}{\Roman{enumi}}
%\renewcommand{\labelenumii}{\arabic{enumii}.}
\renewcommand{\labelenumii}{\Alph{enumii}.}
%\renewcommand{\labelenumiii}{(\alph{enumiii})}
%\renewcommand\thefigure{4.\arabic{figure}}

\begin{description}
\item[Introduction]$\quad$

  \begin{enumerate}
  \item
  \emph{Attention Getter}:
  I'm doing my speech on the legal pot controversy.
  \item
  \emph{Listener Relevance}:
  Just kidding, we aren't allowed to give speeches on topical controversies like that.
  \item
  \emph{Speaker Credibility}:
  Our textbook \cite{text} advises us to speak on many of the ``vital topics'' from our do not speak list.
  \item
  \emph{Thesis}:
  Our speech class could be a valuable tool for making us better and stronger people by allowing us to challenge one another morally, politically, or spiritually.
  \item
  \emph{Preview}:
  Lets talk about how we easily we can insulate ourselves from uncomfortable ideas. Then we can talk about how our university could help make us uncomfortable. Finally we will look at why in the world we would want that.
  \end{enumerate}
\item[Body]$\quad$

\begin{enumerate}
  \item
  \emph{Problem}:
  We live in a world where it is all too easy to insulate ourselves from others who have different points of view. We see the result all around us, grid lock in congress, intolerant billboards, ignorance online, etc.
  \begin{enumerate}
    \item
      Have you ever noticed that the more things you like on Facebook, the less things you see that you don't like?
    \item
      Social media inhibits political and religious discussion \cite{silence}.
    \item
      Political polarization is greater than ever \cite{polar}.
  \end{enumerate}
  \renewcommand{\labelenumi}{\emph{Transition}:}
  \item
  Island Slide
  \renewcommand{\labelenumi}{\Roman{enumi}}
  \setcounter{enumi}{1}
  \item
  \emph{Solution}:
  Our university system is a place where people of all types come together. Instead of washing out and hiding our differences we should shine a light on them. We would strengthen our own understanding of the world and maybe gain some understanding of people who are different.
  \begin{enumerate}
    \item
      According to our bulletin the general education system is supposed to offer ``a dynamic environment for students to participate in improving themselves and the world around them''\cite{bulletin}.
    \item
      If we have never been exposed to challenging ideas, then we can not fully understand our own views.
    \item
      "Honest disagreement is often a good sign of progress."--Mahatma Ghandi \cite{ghandi}
  \end{enumerate}
  \renewcommand{\labelenumi}{\emph{Transition}:}
  \item
  Diversity Slide
  \renewcommand{\labelenumi}{\Roman{enumi}}
  \setcounter{enumi}{2}
  \item
  \emph{Visualization}:
  Imagine graduating, getting a job. On your first day your coworker makes a horrifying comment. What happens next?
  \begin{enumerate}
    \item
      You have been exposed to these ideas before and have already know how to address the situation.
    \item
      You cry. Maybe quit. The job probably won't last that long anyhow \cite{forbes}
    \item
      You cut off their fool head.
  \end{enumerate}
  \renewcommand{\labelenumi}{\emph{Transition}:}
  \item
  Sword Slide
  \renewcommand{\labelenumi}{\Roman{enumi}}
\end{enumerate}
\item[Conclusion]$\quad$
\begin{enumerate}
  \item
  \emph{Thesis Restatement}:
  Our speech class could be a valuable tool for making us better and stronger people by allowing us to challenge one another morally, politically, or spiritually.
  \item
  \emph{Main Point Summary}:
  We have a problem with polarization and isolationism in this country. Our university general education classes would be the perfect place to begin to bridge these gaps. This will make you a better and more resilient citizen/employee/employer.
  \item
  \emph{Clincher}:
  Our beliefs are strong enough to handle challenges and come out stronger, or they are weak enough that they will be destroyed when challenged. Either way, why aren't we challenging them?
\end{enumerate}
\end{description}


\newpage
\bibliographystyle{apacite}
\bibliography{speech-bib-persuasive}
\end{document}
