\documentclass[letterpaper]{article}

\usepackage[cm]{fullpage}
\usepackage{nopageno}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[utf8]{inputenc}
\usepackage{aeguill}
\usepackage{setspace}
\usepackage{apacite}
\usepackage{multicol}
\allowdisplaybreaks

\newcommand{\abs}[1]{\left\lvert #1 \right\rvert}

\begin{document}
\begin{multicols}{2}
%\title{Formal Outline\\ Informative presentation on Évariste Galois}
%\date{March 6, 2015}
%\author{Jon Allen}
%\maketitle
\renewcommand{\labelenumi}{\Roman{enumi}}
%\renewcommand{\labelenumii}{\arabic{enumii}.}
\renewcommand{\labelenumii}{\Alph{enumii}.}
%\renewcommand{\labelenumiii}{(\alph{enumiii})}
%\renewcommand\thefigure{4.\arabic{figure}}

\begin{description}
\item[Introduction]$\quad$

  \begin{enumerate}
  \item
  ``He dies young whom the gods love.'' -- Menander (greek playwright)
  \item
  Major gives perspective.
  
  most interesting is Galois.
  \item
  one of craziest mathies ever.

  expelled, jailed.

  didn't have adulthood.
  \item
  explore life, avoid boring math stuff.
  \item
  historical setting,

  then childhood/early life,
  then death and legacy.
  \end{enumerate}

  citing: about galois: \cite{gods}

  about france: \cite{france}
\item[Body]$\quad$

\begin{enumerate}
  \item
  understand Galois=understand his time: 19th century France.

  \begin{enumerate}
    \item
    1800's=Napoleon

    invested in French universities
    
    peers greatest mathematicians
    \item
    Waterloo 1815/4 after birth.

    France$\in$limbo.

    didn't want king but liked Napoleon.

    $\to$constitutional monarchy.

    king+2 houses, independant courts and press.
    \item
    1824 king dies, new king is brother.

    new king sucks. done by 1830.

    third king not bad, too late.
  \end{enumerate}
  \renewcommand{\labelenumi}{\emph{Transition}:}
  \item
  parents liberal, political. Republicans.

  early influence important.
  \renewcommand{\labelenumi}{\Roman{enumi}}
  \setcounter{enumi}{1}
  \item
  early life and education
  \begin{enumerate}
    \item
    born 1811, Napoleon's son

    father is  mayor Bourg-la-Reine.

    mother strong literary background
    
    home school. turn down college at 10.
    \item
    12 School of Louis the Great.

    awards Latin \& stuff,

    soon bored, original math papers like novel.
    \item
    @17 failed to get into EP.

    didn't prepare

    went to EN (literal).
  \end{enumerate}
  \renewcommand{\labelenumi}{\emph{Transition}:}
  \item
  beginning end.

  starts to make fundamental discoveries
  
  heat up politic.
  \renewcommand{\labelenumi}{\Roman{enumi}}
  \setcounter{enumi}{2}
  \item
  \emph{Third Main Point}: life work and politics
  \begin{enumerate}
    \item
    now attending a 2nd rate school
    
    publishing papers solve centuries old problems
    
    laying down foundations for...algebra.
    \item
    1829 dad suicide priest...
    
    people with his papeprs dying or losing
    \item
    expelled at 19
    
    head of his school a coward in newspaper.

    joins an artillery unit in the National Guard.

    artillery unit disbanded overthrow gvmt.
    
    jailed for conspiracy to overthrow
    \item
    out of jail he threatens kings life

    soon in mysterious duel.  over love/politics.

    night before writes some unpublished ideas.

    \cite{abstract} Hermann Weyl: ``is perhaps the most substantial piece of writing in the whole literature of mankind.''
  \end{enumerate}
  \renewcommand{\labelenumi}{\emph{Transition}:}
  \item
  @20, greatest minds ever dead. genius for liberty.
  \renewcommand{\labelenumi}{\Roman{enumi}}
\end{enumerate}
\item[Conclusion]$\quad$
\begin{enumerate}
  \item
  very briefly at life,

  focussing on color

  only mention brilliance.
  \item
  product times.

  political well educated parents,

  world dominated by intellectuals and politics.

  childhood$\to$death, embodied the best of above.
  \item
  The really great minds are often unnamed. We don't study Greek Government. We study democracy. We don't study Einsteinian Science. We study physics. We don't study Galois Theory. We study algebra.

\begin{quote}
Death is a Dialogue between

The Spirit and the Dust.

"Dissolve" says Death--The Spirit "Sir

I have another Trust"--
\end{quote}--Emily Dickinson
\end{enumerate}
\end{description}
\end{multicols}

\newpage
\bibliographystyle{apacite}
\bibliography{speech-bib-2015-03-06}
\end{document}
