\documentclass{article}
\usepackage{fullpage}
\usepackage{nopageno} 
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{cancel}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, calc}
\allowdisplaybreaks

\newcommand{\abs}[1]{\left\lvert #1 \right\rvert}

\begin{document}
\title{Notes}
\date{April 30, 2014}
\maketitle
counting formula $R_n=\sum\limits_{r=0}^n{\frac{1}{n-r+1}\binom{2n-r}{r,n-r,n-r}}$ where r=flat steps, and $R_n$ is number of large shroder paths.
\begin{align*}
  R_n&=2s_{n+1}\text{ s are dissections/small shrodinger paths}
\end{align*}
\section*{7.1/2}
fibanocci numbers
\begin{align*}
  f_n=f_{n-1}+f_{n-2},f_0=1,f_1=1\\
\end{align*}
partial sums
\begin{align*}
  s_n=f_0+\dots+f_n=f_{n+2}-1
\end{align*}
proof by induction

even odd odd patter

using $f_n$'s to solve covering 2xn chessboard problems.

fibonacci numbersare sum pascal triangle diagonal

generating functions
\begin{align*}
  g(x)&=\sum\limits_{n=0}^\infty{h_nx^n}
\end{align*}

derive recurrence, use recurrence
use reg generating functions for combination like things
\section*{7.3}
exponential generating functions
use for counting permutation like things.
need to be able to find a generating function and use it to find a counting formula.
\section*{7.4/5}
solving recurrence relations and ....

know if it's linear, homogeneous, constant coefficients.

method of generating functions know this.

also can find general solution to homogen, find particular solution, combine solutions and determince constants.
\section*{7.6/8.1}
catalan stuff.
catalan number is $C_n=\frac{1}{n+1}\binom{2n}{n}$

say what is counted, give bijections. say why it's a bijection.
\section*{8.2}

\section*{8.3}
self conjugate partitions, and partitions with only odd parts.

odd and distinct
\section*{8.5}
\end{document}
